---
title: "p8105_hw3_by2287"
author: "Binyam Yilma"
output: github_document
---

#### Setup

```{r, include = FALSE}
library(tidyverse)
library(readr)

library(p8105.datasets)
data("instacart")

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6, 
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
 ggplot2.continous.colour = "viridis",
 ggplot2.continous.fill = "viridis"
)
scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```



The dataframe `instacart` is made up of `r nrow(instacart)` rows and `r ncol(instacart)` columns, and it contains information on grocery orders from  `r instacart %>% pull(user_id)  %>% n_distinct()` unique customers.
There are `r instacart %>% pull(user_id)  %>% n_distinct()` unique products in this dataset. 

Some key variables in `instacart` include: the `order_id` for a particular order, and an identifier for the each customer in `user_id`, the names of different products in `product_name`, as well as the `order_hour_of_day`, which has a range of: `r instacart %>% pull(order_hour_of_day) %>% range()`. 

How many aisles are there, and which aisles are the most items ordered from?

There are `r instacart %>% pull(aisle_id)  %>% n_distinct()` unique aisles. 

These are the top 3 most ordered from aisles. 
```{r, echo = FALSE, message  = F}
instacart %>% 
  group_by(aisle) %>% 
  summarise(n_orders = n()) %>% 
  arrange(desc(n_orders)) %>% 
  head(3) %>% 
  knitr::kable()
```


A plot that shows the number of items ordered in each aisle (looking at only aisles with > 10000 orders), limiting this to aisles with more than 10000 items ordered. 

```{r, message = FALSE}
instacart %>% 
  group_by(aisle) %>% 
  summarise(n_orders = n()) %>% 
  arrange(desc(n_orders)) %>% 
  filter(n_orders > 10000) %>% 
  ggplot(aes(x = aisle, y = n_orders)) + 
  geom_point() + 
  theme(axis.text.x = element_text(angle = 60, vjust = 1.0, hjust = 1))

```

The three most popular items in the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”:

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(
    rank = min_rank(desc(n))
      ) %>% 
  filter(rank %in% c(1,2,3)) %>% 
  arrange(aisle, rank) %>% 
  knitr::kable()
```


The mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week:

```{r, message = FALSE}
instacart %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(product_name, order_dow) %>% 
  summarise(mean_hour = mean(order_hour_of_day)) %>% 
  pivot_wider(
		names_from = order_dow,
		values_from = mean_hour
	) %>% 
  knitr::kable(digits = 2)
```



Problem 2

Load, tidy, and otherwise wrangle the data. Your final dataset should include all originally observed variables and values; have useful variable names; include a weekday vs weekend variable; and encode data with reasonable variable classes. Describe the resulting dataset (e.g. what variables exist, how many observations, etc).
Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate accross minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?
Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.

## Problem 2 
```{r}

accel = read_csv("./data/accel_data.csv") %>% 
  janitor::clean_names() 

accel_tidy = accel %>%
  pivot_longer(
    cols = activity_1:activity_1440, 
    names_to = "activity_id", 
    names_prefix = "activity_",
    values_to = "activity_count"
    ) %>% 
  mutate(
    day_type = case_when(
      day %in% c("Saturday","Sunday") ~ "Weekend",
      TRUE ~ "Weekday"
    ), 
    day_id = as.integer(day_id),
    week = as.integer(week),
    activity_id = as.integer(activity_id),
    day = factor(day)
  ) %>% 
  relocate(day_id, day, day_type, week)
```


The data frame `accel_tidy` is a tidy version of the `accel` dataset, imported in the above code chunk. 

`accel_tidy` has `r nrow(accel_tidy)` rows and `r ncol(accel_tidy)` columns, and it includes these variables: `r accel_tidy %>% names`. 

These variables are collected from an accelerometer, which was worn by a 63 year-old male, and contain minute by minute information on `r accel_tidy %>% pull(activity_id) %>% n_distinct()` different types of activities, measured for each day over 5 weeks. 






Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate accross minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?



#### total activity over a given day
```{r}
accel_tidy %>% 
  group_by(day_id, day, day_type) %>% 
  summarise(
    total_activity = sum(activity_count)
  ) %>% 
  arrange(day_id) %>% 
  knitr::kable(digits = 2)
  # ggplot(aes(x = day_id, y = total_activity)) + 
  # geom_point(alpha = .5) + 
  # geom_smooth(se = F) + 
  # theme(axis.text.x = element_text(angle = 60, vjust = 1.0, hjust = 1))
```

It's hard to spot any trends in the data, in specific the total activity per day, from the table only. 

Worth noting here, is that there a couple of Saturday's where  the subject's activity level was exactly 1440.0 (perhaps an artifact of the accelerometer?).

#### Graph showing the 24-hour activity time courses for each day
```{r}
accel_tidy %>% 
  ggplot(aes(y = day_id, x = activity_count, color = day)) + 
  geom_point(alpha = .5) + 
  scale_y_continuous(
    breaks = c(1:35)
    ) 
```

